Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-03-18 ltx/BrainIDE.tex
further updated paper following meeting on the 20/03/2010

% &&\mathbf{\theta}_1\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_1^2}\right)}+\mathbf{\theta}_2\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_2^2}\right)}+\mathbf{\theta}_3\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_3^2}\right)}\nonumber

@@

% % 	\mathbf{\theta}=\begin{bmatrix} \mathbf{\theta}_{1}\\\mathbf{\theta}_{2}\\\mathbf{\theta}_{3}\end{bmatrix},\quad 		\Sigma_w=\begin{bmatrix}\Sigma_{w1}\\\Sigma_{w2}\\ \Sigma_{w3}\end{bmatrix}.

@@

is a spatially colored, temporally white iid. This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 

@@

	y_t(\mathbf{r}_n) = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \varepsilon_t

@@

% \begin{equation}
%  cov(e_t(\mathbf{r}),e_{t+\tau}(\mathbf{r'}))=
% \begin{cases}
% \beta\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
% 0 & \mathrm{otherwise}
% \end{cases}
% \label{eq:FieldCovariance}
% \end{equation}
% \section{Generating Data}
% To generate iEEG/LFP data we proceeded with the assumption of spatial homogeneity for the connectivity and activation function parameters, implying $w(\mathbf{r},\mathbf{r}') = w\left(\mathbf{r}-\mathbf{r}'\right)$, $\nu_0(\mathbf{r}) = \nu_0$ and $v_0(\mathbf{r}) = v_0$. In addition we assume the synaptic transmission delay is negligible. The model was discretized in time   and a stochastic 
% 
% The IDE formulation in Eq.~\ref{DiscreteTimeModel} was solved using a numerical method described in Appendix~\ref{Space Discretization}. 
% All the model parameters are described in Table 1.
% 
% \begin{tabular}{|c|c|c|}
% \hline
% Parameter & Value & Reference\\
% \hline
% $\alpha$ & $100$ s$^{-1}$ &\\
% \hline
% $\nu_0$ & $1$ s$^{-1}$ &\\
% \hline
% $v_0$ & $6$ mV &\\
% \hline
% $\beta$ & $0.56$ spikes/mV$^{-1}$ & \\
% \hline
% $T$ & $1\times10^{-4}$ s &\\
% \hline
% $\Delta$ & $1$ &\\
% \hline
% $\mathbf{\theta}_1$ &  &\\
% \hline
% $\mathbf{\theta}_2$ &  &\\
% \hline
% $\mathbf{\theta}_3$ &  &\\
% \hline
% $\sigma_{w1}$ &  &\\
% \hline
% $\sigma_{w2}$ &  &\\
% \hline
% $\sigma_{w3}$ &  &\\
% \hline
% $\sigma_m$ &  &\\
% \hline
% \end{tabular}

\section{Derivation of Finite Dimensional State-Space Model}
In order to implement standard estimation techniques we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
	v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t
\end{equation}
where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. As described in Eq.~\ref{Connectivity Kernel} the connectivity kernel can also be considered a decomposition into a set of basis functions in 
\begin{equation}\label{DefKernelDecomp}
	w\left(\mathbf{r}-\mathbf{r}'\right) = \mathbf{\theta}^{\top} \mathbf{\psi}\left(\mathbf{r}-\mathbf{r}'\right)
\end{equation}
We will assume we know the parametric form of the connectivity basis functions, where the parameter $\mathbf{\theta}$ is unknown. Each connectivity basis function can be considered a layer in the Wilson and Cowan model, representing short range excitation, surround inhibition and mid-range excitation. Making substitutions of~\ref{DefFieldDecomp} and~\ref{DefKernelDecomp} into~\ref{DiscreteTimeModel} we get
\begin{equation}\label{reduced continuous model}
	\boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T\mathbf{\theta}^{\top}\int_\Omega  {\psi \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'}  - \lambda\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right).
\end{equation}
By defining the term
\begin{equation}\label{DefGamma}
	\Gamma  = \int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)d\mathbf{r}}
\end{equation}
and manipulating the model (see Appendix~\ref{Simplifying Decomposition}) we isolate the state vector to form a state-space model
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_t &=& q(\mathbf{x}_t,\mathbf{\theta},\lambda) + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	q(\mathbf{x}_t,\mathbf{\theta},\alpha) &=& T\Gamma^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\psi^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \mathbf{\theta} - \lambda \mathbf{x}_t \\
	k(u_t(\mathbf{r})) &=& \Gamma^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} \\
	\mathbf{e}_t &=& \Gamma^{-1}\int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}} \\
\end{eqnarray}
where $\mathbf{e}_t$ is a zero-mean normal white noise term with the covariance (see Appendix D)
\begin{equation} 
\mathbf \Sigma_w=\mathbf{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \beta\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\mathbf{\Gamma}^{- \top} 
\end{equation}

where $\tilde{e}_t \sim \mathcal{N}\left(0,R' \Sigma R\right)$ and $R$ is the Cholesky decomposition of $\Gamma^{-1}$. The observation equation of the reduced model is
\begin{eqnarray}\label{ObservationEquation}
    y_t &=& z\left(x_t\right) + \varepsilon_t \\
	z\left(x_t\right) &=& \int_\Omega m\left(|r_n-r|\right)\boldsymbol{\phi}^{\top}\left(r\right) dr x_t.
\end{eqnarray}

@@

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	x_t = q(x_t,\mathbf{\theta},\lambda) + k(I_t) + \tilde{e}_t \\
	y_t = z(x_t) + \varepsilon_t,
\end{eqnarray}
where $\mathbf{\theta}$ and $\lambda$ are the systems parameters.

@@

\section{Numerical Solution to IDE Model}\label{Space Discretization}
To solve the intregro-difference equation in Eq.~\ref{Discrete Time Model1} we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v_{t+1}\left(r_{ij}\right) =  \lambda v_t(r_{ij}) + T \Delta^2 \sum_{i=1}^{i=I}{\sum_{j=1}^{j=J}{\tilde{w}\left( |r_{ij}-r_{ij}'| \right)f\left( v_t\left( r_{ij}'\right) \right)} } + T I_t\left(r\right) + \tilde{e}_t(r_{i,j})
\end{equation}
where $\tilde{e}_t(r_{i,j}) \sim \mathcal{N}\left(0,\Sigma\right)$ and $\tilde{w}(|r_{ij}-r_{ij}'|)$ discretized connectivity kernel described by
\begin{equation}\label{Discrete Connectivity Kernel}
	\tilde{w}(r_{ij},r_{ij}') = \mathbf{\theta}^{\top} e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
\end{equation}
The observation equation and kernel were discretized in the same way by
\begin{equation}\label{Discrete Observation Equation}
    y_k\left(r_{i_n j_n}\right) = \Delta^2\sum_{i=1}^{i=I}\sum_{j=1}^{j=J} \tilde{m}( |r_{i_n j_n}-r_{ij}'| ) v_k\left(r_{ij}'\right)  + \varepsilon_k( r_{i_n j_n})
\end{equation}
where
\begin{equation}\label{Discrete Observation Kernel}
	\tilde{m}(|r_{ij}-r_{ij}'|) = e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
\end{equation}
We have used periodic boundary conditions to alleviate problems associated with edge effects.

@@

	\int_\Omega  {\boldsymbol{\phi} \left(r\right)\boldsymbol{\phi}^{\top}\left(r\right) dr} x_{t+1} = T \int_\Omega  {\boldsymbol{\phi} (r) \mathbf{\theta}^{\top} \int_\Omega  {\psi \left(|r-r'|\right)f\left(\boldsymbol{\phi}^{\top}\left(r'\right) x_t \right)dr'}dr}  - \lambda\int_\Omega {\boldsymbol{\phi}(r)\boldsymbol{\phi}^{\top}(r)dr} x_t + T \int_\Omega{\boldsymbol{\phi} \left(r\right) I_t\left(r\right)dr} + \int_\Omega{\boldsymbol{\phi} \left(r\right) e_t\left(r\right)dr}.

@@

	x_{t+1} = T\Gamma^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(r\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(r'\right)x_t\right)\psi^{\top} \left(|r-r'|\right)dr'} dr} \mathbf{\theta} - \lambda x_t + \Gamma^{-1}T \int_\Omega{\boldsymbol{\phi} \left(r\right) I_t\left(r\right)dr} + \Gamma^{-1}\int_\Omega{\boldsymbol{\phi}\left(r\right)e_t\left(r\right)dr}.

@@

\begin{equation}
cov(e_t(\mathbf{r}),e_{t+\tau}(\mathbf{r'}))=
\begin{cases}
\beta\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
0 & \mathrm{otherwise}
\end{cases}
\label{eq:FieldCovariance}
\end{equation}
The final state-space model is given by
\begin{equation}
x_{t + 1} = T q(x_t)\mathbf{\theta} -\lambda x_t + \Gamma^{-1}\int_\Omega  {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )d\mathbf{r}}
\end{equation}
then the covariance of the noise term is given by
\begin{eqnarray}
\mathbf{\Sigma}_w&=&\mathbf{\Gamma}^{-1}\mathbf E[\int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r} \int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top} e_t\left(\mathbf{r}'\right)d\mathbf{r}']\mathbf{\Gamma}^{- \top} \nonumber \\
&=&\mathbf{\Gamma}^{-1}\iint\limits_{\Omega} \boldsymbol{\phi}\left(\mathbf{r}\right) \mathbf E[e_t\left(\mathbf{r}\right)e_t\left(\mathbf{r}'\right)]\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf r\mathbf{\Gamma}^{- \top} \nonumber\\
&=&\mathbf{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \beta\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 
\end{eqnarray}

\section{Parameter estimation using LS}
\begin{equation}
 \mathbf x_{t+1}=\mathbf{x}_{t}+T_s \mathbf{Q}(\mathbf{x}_t)\mathbf \mathbf{\theta}-\alpha T_s\mathbf{x}_t+T_s\Gamma^{-1}\int_\Omega\boldsymbol{\phi}(\mathbf r)\varepsilon_t(\mathbf r)d\mathbf r
\label{eq:statespacemodel}
\end{equation}
where the matrix $\mathbf Q_{n_x \times n_{\mathbf{\theta}}}(\mathbf x_t)$ is defined as
\begin{equation}
 Q_{n_x \times n_{\mathbf{\theta}}}=\Gamma^{-1}\int_\Omega\boldsymbol{\phi}(\mathbf r)\int_\Omega f(\boldsymbol{\phi}^\top(\mathbf r')\mathbf x_t)\mathbf{\psi}^\top(\mathbf r-\mathbf r')d\mathbf r'd\mathbf r
\end{equation}
and $T_s$ is the sampling period. Least Square method can be used as (\ref{eq:statespacemodel}) is linear in parameters. At each time instant we have

>>>>

% &&\boldsymbol{\theta}_1\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_1^2}\right)}+\boldsymbol{\theta}_2\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_2^2}\right)}+\boldsymbol{\theta}_3\exp{\left(-\frac{(r-r')^\top(r-r')}{\sigma_3^2}\right)}\nonumber

@@

% % 	\boldsymbol{\theta}=\begin{bmatrix} \boldsymbol{\theta}_{1}\\\boldsymbol{\theta}_{2}\\\boldsymbol{\theta}_{3}\end{bmatrix},\quad 		\Sigma_w=\begin{bmatrix}\Sigma_{w1}\\\Sigma_{w2}\\ \Sigma_{w3}\end{bmatrix}.

@@

is an iid spatially colored, temporally white disturbance term. This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 

@@

	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t

@@

% \section{Generating Data}
% To generate iEEG/LFP data we proceeded with the assumption of spatial homogeneity for the connectivity and activation function parameters, implying $w(\mathbf{r},\mathbf{r}') = w\left(\mathbf{r}-\mathbf{r}'\right)$, $\nu_0(\mathbf{r}) = \nu_0$ and $v_0(\mathbf{r}) = v_0$. In addition we assume the synaptic transmission delay is negligible. The model was discretized in time   and a stochastic 
% 
% The IDE formulation in Eq.~\ref{DiscreteTimeModel} was solved using a numerical method described in Appendix~\ref{Space Discretization}. 
% All the model parameters are described in Table 1.
% 
% \begin{tabular}{|c|c|c|}
% \hline
% Parameter & Value & Reference\\
% \hline
% $\alpha$ & $100$ s$^{-1}$ &\\
% \hline
% $\nu_0$ & $1$ s$^{-1}$ &\\
% \hline
% $v_0$ & $6$ mV &\\
% \hline
% $\beta$ & $0.56$ spikes/mV$^{-1}$ & \\
% \hline
% $T$ & $1\times10^{-4}$ s &\\
% \hline
% $\Delta$ & $1$ &\\
% \hline
% $\mathbf{\theta}_1$ &  &\\
% \hline
% $\mathbf{\theta}_2$ &  &\\
% \hline
% $\mathbf{\theta}_3$ &  &\\
% \hline
% $\sigma_{w1}$ &  &\\
% \hline
% $\sigma_{w2}$ &  &\\
% \hline
% $\sigma_{w3}$ &  &\\
% \hline
% $\sigma_m$ &  &\\
% \hline
% \end{tabular}

\section{Derivation of Finite Dimensional State-Space Model}
In order to implement standard estimation techniques we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
	v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t
\end{equation}
where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. As described in Eq.~\ref{Connectivity Kernel} the connectivity kernel can also be considered a decomposition into a set of basis functions in 
\begin{equation}\label{DefKernelDecomp}
	w\left(\mathbf{r}-\mathbf{r}'\right) = \boldsymbol{\theta}^{\top} \mathbf{\psi}\left(\mathbf{r}-\mathbf{r}'\right)
\end{equation}
We will assume we know the parametric form of the connectivity basis functions, where the parameter $\boldsymbol{\theta}$ is unknown. Each connectivity basis function can be considered a layer in the Wilson and Cowan model, representing short range excitation, surround inhibition and mid-range excitation. Making substitutions of~\ref{DefFieldDecomp} and~\ref{DefKernelDecomp} into~\ref{DiscreteTimeModel} we get
\begin{equation}\label{reduced continuous model}
	\boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T\boldsymbol{\theta}^{\top}\int_\Omega  {\psi \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'}  - \lambda\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right).
\end{equation}
By defining the term
\begin{equation}\label{DefGamma}
	\boldsymbol{\Gamma}  = \int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)d\mathbf{r}}
\end{equation}
and manipulating the model (see Appendix~\ref{Simplifying Decomposition}) we isolate the state vector to form a state-space model
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_t &=& q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) &=& T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\psi^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\
	k(u_t(\mathbf{r})) &=& \boldsymbol{\Gamma}^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} \\
	\mathbf{e}_t &=& \boldsymbol{\Gamma}^{-1}\int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}} \\
\end{eqnarray}
where $\mathbf{e}_t$ is a zero-mean normal white noise term with the covariance (see Appendix D)
\begin{equation} 
\boldsymbol{\Sigma}_w=\boldsymbol{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \rho\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 
\end{equation}

\begin{equation}
 cov(\mathbf{e}_t,\mathbf{e}_{t+\tau})=
\begin{cases}
\rho\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
0 & \mathrm{otherwise}
\end{cases}
\label{eq:FieldCovariance}
\end{equation}

The observation equation of the reduced model is
\begin{eqnarray}\label{ObservationEquation}
    \mathbf{y}_t &=& z\left(\mathbf{x}_t\right) + \boldsymbol{\varepsilon}_t \\
	z\left(\mathbf{x}_t\right) &=& \int_\Omega m\left(\mathbf{r}_n-\mathbf{r}'\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}' \mathbf{x}_t.
\end{eqnarray}

@@

The approximation of the state density is made by propagating the so called sigma points through the state-space model. The sigma points describe the mean and covariance of the state estimate. The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = z(\mathbf{x}_t) + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\alpha$ are the systems parameters.

@@

\section{Numerical Solution to IDE Model}\label{Space Discretization}
To solve the intregro-difference equation in Eq.~\ref{Discrete Time Model1} we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v_{t+1}\left(r_{ij}\right) =  \lambda v_t(r_{ij}) + T \Delta^2 \sum_{i=1}^{i=I}{\sum_{j=1}^{j=J}{\tilde{w}\left( |r_{ij}-r_{ij}'| \right)f\left( v_t\left( r_{ij}'\right) \right)} } + T I_t\left(r\right) + \tilde{e}_t(r_{i,j})
\end{equation}
where $\tilde{e}_t(r_{i,j}) \sim \mathcal{N}\left(0,\Sigma\right)$ and $\tilde{w}(|r_{ij}-r_{ij}'|)$ discretized connectivity kernel described by
\begin{equation}\label{Discrete Connectivity Kernel}
	\tilde{w}(r_{ij},r_{ij}') = \boldsymbol{\theta}^{\top} e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
\end{equation}
The observation equation and kernel were discretized in the same way by
\begin{equation}\label{Discrete Observation Equation}
    y_k\left(r_{i_n j_n}\right) = \Delta^2\sum_{i=1}^{i=I}\sum_{j=1}^{j=J} \tilde{m}( |r_{i_n j_n}-r_{ij}'| ) v_k\left(r_{ij}'\right)  + \varepsilon_k( r_{i_n j_n})
\end{equation}
where
\begin{equation}\label{Discrete Observation Kernel}
	\tilde{m}(|r_{ij}-r_{ij}'|) = e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
\end{equation}
We have used periodic boundary conditions to alleviate problems associated with edge effects.

@@

	\int_\Omega  {\boldsymbol{\phi} \left(r\right)\boldsymbol{\phi}^{\top}\left(r\right) dr} x_{t+1} = T \int_\Omega  {\boldsymbol{\phi} (r) \boldsymbol{\theta}^{\top} \int_\Omega  {\psi \left(|r-r'|\right)f\left(\boldsymbol{\phi}^{\top}\left(r'\right) x_t \right)dr'}dr}  - \lambda\int_\Omega {\boldsymbol{\phi}(r)\boldsymbol{\phi}^{\top}(r)dr} x_t + T \int_\Omega{\boldsymbol{\phi} \left(r\right) I_t\left(r\right)dr} + \int_\Omega{\boldsymbol{\phi} \left(r\right) e_t\left(r\right)dr}.

@@

	x_{t+1} = T\Gamma^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(r\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(r'\right)x_t\right)\psi^{\top} \left(|r-r'|\right)dr'} dr} \boldsymbol{\theta} - \lambda x_t + \Gamma^{-1}T \int_\Omega{\boldsymbol{\phi} \left(r\right) u_t\left(r\right)dr} + \Gamma^{-1}\int_\Omega{\boldsymbol{\phi}\left(r\right)e_t\left(r\right)dr}.

@@

\begin{equation}
cov(e_t(\mathbf{r}),e_{t+\tau}(\mathbf{r'}))=
\begin{cases}
\beta\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
0 & \mathrm{otherwise}
\end{cases}
\label{eq:FieldCovariance}
\end{equation}
The final state-space model is given by
\begin{equation}
x_{t + 1} = T q(x_t)\boldsymbol{\theta} -\lambda x_t + \Gamma^{-1}\int_\Omega  {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )d\mathbf{r}}
\end{equation}
then the covariance of the noise term is given by
\begin{eqnarray}
\mathbf{\Sigma}_w&=&\mathbf{\Gamma}^{-1}\mathbf E[\int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r} \int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top} e_t\left(\mathbf{r}'\right)d\mathbf{r}']\mathbf{\Gamma}^{- \top} \nonumber \\
&=&\mathbf{\Gamma}^{-1}\iint\limits_{\Omega} \boldsymbol{\phi}\left(\mathbf{r}\right) \mathbf E[e_t\left(\mathbf{r}\right)e_t\left(\mathbf{r}'\right)]\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf r\mathbf{\Gamma}^{- \top} \nonumber\\
&=&\mathbf{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \beta\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 
\end{eqnarray}

\section{Parameter estimation using LS}
\begin{equation}
 \mathbf x_{t+1}=\mathbf{x}_{t}+T_s \mathbf{Q}(\mathbf{x}_t)\boldsymbol{\theta}-\alpha T_s\mathbf{x}_t+T_s\Gamma^{-1}\int_\Omega\boldsymbol{\phi}(\mathbf r)\varepsilon_t(\mathbf r)d\mathbf r
\label{eq:statespacemodel}
\end{equation}
where the matrix $\mathbf Q_{n_x \times n_{\boldsymbol{\theta}}}(\mathbf x_t)$ is defined as
\begin{equation}
 Q_{n_x \times n_{\boldsymbol{\theta}}}=\Gamma^{-1}\int_\Omega\boldsymbol{\phi}(\mathbf r)\int_\Omega f(\boldsymbol{\phi}^\top(\mathbf r')\mathbf x_t)\mathbf{\psi}^\top(\mathbf r-\mathbf r')d\mathbf r'd\mathbf r
\end{equation}
and $T_s$ is the sampling period. Least Square method can be used as (\ref{eq:statespacemodel}) is linear in parameters. At each time instant we have

\begin{eqnarray}
 \mathbf x_{1}&=&\mathbf x_{0}+T_s \mathbf Q(\mathbf x_0) \boldsymbol{\theta}-\alpha T_s\mathbf x_0+T_s\mathbf e_0 \nonumber \\
 \mathbf x_{2}&=&\mathbf x_{1}+T_s \mathbf Q(\mathbf x_1) \boldsymbol{\theta}-\alpha T_s\mathbf x_1+T_s\mathbf e_1\nonumber\\
&\vdots& \nonumber\\
 \mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1}) \boldsymbol{\theta}-\alpha T_s\mathbf x_{n-1}+T_s\mathbf e_{n-1}
\end{eqnarray}
in matrix form
\begin{equation}
 \begin{bmatrix} \mathbf x_{1}\\\mathbf x_{2}\\\vdots\\\mathbf x_{n}\end{bmatrix}-\begin{bmatrix} \mathbf x_{0}\\\mathbf x_{1}\\\vdots\\\mathbf x_{n-1}\end{bmatrix}=T_s\begin{bmatrix}\mathbf Q(\mathbf x_0)&-\mathbf x_{0}\\\mathbf Q(\mathbf x_1)&-\mathbf x_{1}\\\vdots\\ \mathbf Q(\mathbf x_{n-1})&-\mathbf x_{n-1}\end{bmatrix}\begin{bmatrix} \boldsymbol{\theta} \\ \alpha\end{bmatrix}+T_s\begin{bmatrix}\mathbf e_0\\\mathbf e_1\\\vdots\\\mathbf e_{n-1}\end{bmatrix}
\end{equation}
or in compact form we have
\begin{equation}
 \mathbf Z=\mathbf X \mathbf W+\boldsymbol \xi
\end{equation}
where
\begin{equation}
 \mathbf Z=\begin{bmatrix} \mathbf x_{1}-\mathbf x_{0}\\\mathbf x_{2}-\mathbf x_{1}\\\vdots\\\mathbf x_{n}-\mathbf x_{n-1}\end{bmatrix},\quad \mathbf X=T_s\begin{bmatrix}\mathbf Q(\mathbf x_0)&-\mathbf x_{0}\\\mathbf Q(\mathbf x_1)&-\mathbf x_{1}\\\vdots\\ \mathbf Q(\mathbf x_{n-1})&-\mathbf x_{n-1}\end{bmatrix},\quad \mathbf W=\begin{bmatrix} \boldsymbol{\theta} \\ \alpha\end{bmatrix} \text{and}\quad \boldsymbol \xi=T_s\begin{bmatrix}\mathbf e_0\\\mathbf e_1\\\vdots\\\mathbf e_{n-1}\end{bmatrix}
\end{equation}
using the method of Least Square $ \mathbf W$ can be estimated
\begin{equation}
 \mathbf W\approx(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z
\end{equation}
% use section* for acknowledgement
\section{clarification for the derivation} 
I ignore noise here. From equation (\ref{SpikesToPotential}) we can say $h$ is the foundamental solution to the equation
\begin{equation}
 Lv(r,t)=g(r,t)
\label{eq:LV}
\end{equation}
where $L$ is a temporal differential operator, since $h(t)$ is the foundamental solution to (\ref{eq:LV}) it can be written as a Green's function
\begin{equation}
 Lh(t)=\delta(t)
\end{equation}
where $L=1+\tau\frac{d}{dt}$ therefore equation (\ref{eq:LV}) becomes:
\begin{equation}
\tau\frac{dv(r,t)}{dt}+ v(r,t)=g(r,t)
\end{equation}
substituting (\ref{RateBasedInteractions}) for $g(r,t)$ gives
\begin{equation}
\frac{dv\left( {r,t} \right)}{dt} + \frac{1}{\tau} v\left( r,t \right) = \frac{1}{\tau}\int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'} 
\end{equation}
